### 2023-03-24 example.Rmd
barplot

---
title: "An example of extremely reproducible enrichment analysis"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/enrichment_recipe

## Introduction

This guide is a Rmarkdown script that conducts differential expression and enrichment analysis,
which are very popular workflows for transcriptome data.

This is meant to be a boilerplate template, which you can remix and modify to suit your analytical
needs.

In the code chunk below called `libs`, you can add and remove required R library dependancies.
Check that the libraries listed here match the Dockerfile, otherwise you might get errors.

```{r,libs}

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("clusterProfiler")
  library("fgsea")
  library("eulerr")
  library("gplots")
})

```

For this guide I will be using bulk RNA-seq data from a previous study, which is
deposited at NCBI GEO and SRA under accession numbers: GSE55123 and SRP038101 (Lund et al, 2014).
The experiment is designed to investigate the effect of Azacitidine treatment on AML3 cells.

The raw data have been processed by the DEE2 project, and the summary gene expression counts are
available at the dee2.io website, and programmatically with the getDEE2 bioconductor package
(Ziemann et al, 2019).

Alternatively, you could fetch data from another resource like NCBI GEO, Zenodo or from the host
storage drive.

```{r,fetch}

mdat <- getDEE2Metadata("hsapiens")

# get sample sheet
ss <- subset(mdat,SRP_accession=="SRP038101")

# fetch the whole set of RNA-seq data
x <- getDEE2("hsapiens",ss$SRR_accession , metadata=mdat, legacy=TRUE)
mx <- x$GeneCounts
rownames(mx) <- paste(rownames(mx),x$GeneInfo$GeneSymbol)
dim(mx)

# aza no filtering
ss$trt <- grepl("Treated",ss$Experiment_title)

ss %>%
  kbl(caption="sample sheet for Aza treatment in AML3 cells") %>%
  kable_paper("hover", full_width = F)

```

## Data quality control

QC is important, even if you are using public transcriptome data.
For RNA-seq it is a good idea to show the number of reads for each sample.

```{r,counts1}

par(mar=c(5,7,5,1))
barplot(rev(colSums(mx)),horiz=TRUE,las=1,main="number of reads per sample in SRP038101")

```

## Differential expression analysis

I will be using DESeq2 for DE analysis, the count matrix is prefiltered using
a detection threshold of 10 reads per sample across all samples.
All genes that meet the detection threshold will comprise the background list.

```{r,de1}

mxf <- mx[which(rowMeans(mx)>=10),]
dim(mxf)
dds <- DESeqDataSetFromMatrix(countData = mxf , colData = ss, design = ~ trt )
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <-cbind(as.data.frame(z),mxf)
def <-as.data.frame(zz[order(zz$pvalue),])

head(def,10) %>%
  kbl(caption="Top DE genes for Aza treatment") %>%
  kable_paper("hover", full_width = F)

sigf <- subset(def,padj<=0.05)

DET=nrow(mxf)
NSIG=nrow(sigf)
NUP=nrow(subset(sigf,log2FoldChange>0))
NDN=nrow(subset(sigf,log2FoldChange<0))

HEADER=paste(DET,"detected genes;",NSIG,"w/FDR<0.05;",NUP,"up;",NDN,"down")

plot(log10(def$baseMean) ,def$log2FoldChange,
  cex=0.6,pch=19,col="darkgray",
  main="5-azacitidine treatment in AML3 cells",
  xlab="log10(basemean)",ylab="log2(fold change)")

points(log10(sigf$baseMean) ,sigf$log2FoldChange,
  cex=0.6,pch=19,col="red")

mtext(HEADER)

defup <- rownames(subset(def,padj<=0.05 & log2FoldChange>0))
defup <- unique(sapply(strsplit(defup," "),"[[",2))

defdn <- rownames(subset(def,padj<=0.05 & log2FoldChange<0))
defdn <- unique(sapply(strsplit(defdn," "),"[[",2))

bg <- rownames(def)
bg <- unique(sapply(strsplit(bg," "),"[[",2))

```

In the next sections I will run enrichment analysis with over-representation test
and compare it to functional class scoring.

## Enrichment using over-representation test

I've compiled a reporting checklist:

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 |

```{r,gs}
# from https://reactome.org/download/current/ReactomePathways.gmt.zip
genesets2 <- read.gmt("ReactomePathways_2023-03-06.gmt")
gsets <- gmtPathways("ReactomePathways_2023-03-06.gmt")

```

Firstly with the upregulated genes

```{r,oraup}

ora_up <- as.data.frame(enricher(gene = defup ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora_up$geneID <- NULL
ora_up <- subset(ora_up,p.adjust<0.05 & Count >=10)
ora_ups <- rownames(ora_up)

gr <- as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",2))

ora_up$es <- gr/br
ora_up <- ora_up[order(-ora_up$es),]
ora_up$Description=NULL

head(ora_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

topup2 <- rev(head(ora_up$es,10))
names(topup2) <- rev(head(ora_up$ID,10))

```

Now with the downregulated genes

```{r,oradn}

ora_dn <- as.data.frame(enricher(gene = defdn ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora_dn$geneID <- NULL
ora_dn <- subset(ora_dn,p.adjust<0.05 & Count >=10)
ora_dns <- rownames(ora_dn)

gr <- as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",2))

ora_dn$es <- gr/br
ora_dn <- ora_dn[order(-ora_dn$es),]
ora_dn$Description=NULL

head(ora_dn) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

topdn2 <- head(ora_dn$es,10)
names(topdn2) <- head(ora_dn$ID,10)

```

Make a barplot

```{r,topbarplot2,fig.height=6}

par(mar=c(5,20,5,1))

cols <- c(rep("blue",20),rep("red",20))

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE Reactomes",
  xlab="ES")

mtext("ORA test")

```

## Enrichment analysis with functional class scoring

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | FGSEA (check version at foot of report) |
| Statistical test used | Kolmogorov-Smirnov test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 |

```{r,fgsea}

def2 <- def
def2$genename <- sapply(strsplit(rownames(def2)," "),"[[",2)
def2 <- def2[,c("stat","genename")]
def2 <- aggregate(. ~ genename, def2, sum)
stat <- def2$stat
names(stat) <- def2$genename
fgseaRes <- fgsea(pathways=gsets, stats=stat, minSize=10, nPermSimple = 10000)
fgseaRes <- fgseaRes[order(fgseaRes$pval),]

```

```{r,fgsea_sig}

fgsea_up <- subset(fgseaRes,padj<0.05 & ES>0)
fgsea_ups <- fgsea_up$pathway
fgsea_dn <- subset(fgseaRes,padj<0.05 & ES<0)
fgsea_dns <- fgsea_dn$pathway

fgsea_up <-  data.frame(fgsea_up[order(-fgsea_up$ES),])

fgsea_dn <-  data.frame(fgsea_dn[order(fgsea_dn$ES),])

head(fgsea_up,10) %>%
  kbl(row.names = FALSE, caption="FGSEA:Top upregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

head(fgsea_dn,10) %>%
  kbl(row.names = FALSE, caption="FGSEA:Top downregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

fgsea_up <- head(fgsea_up,20)
fgsea_up_vec <- fgsea_up$ES
names(fgsea_up_vec) <- fgsea_up$pathway
fgsea_up_vec <- rev(fgsea_up_vec)

fgsea_dn <- head(fgsea_dn,20)
fgsea_dn_vec <- fgsea_dn$ES * -1
names(fgsea_dn_vec) <- fgsea_dn$pathway

```

Make a barplot

```{r,topbarplot3,fig.height=6}

par(mar=c(5,20,5,1))

cols <- c(rep("blue",20),rep("red",20))

barplot(c(fgsea_dn_vec,fgsea_up_vec),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE Reactomes",
  xlab="ES")

mtext("FCS test")

```

## Compare pathway results

```{r,venn1}

v1 <- list("ORA up"=ora_ups, "FCS up"=fgsea_ups)
v2 <- list("ORA dn"=ora_dns, "FCS dn"=fgsea_dns)
par(mar=c(10,10,10,10))
par(mfrow=c(2,1))
plot(euler(v1),quantities = TRUE)
plot(euler(v2),quantities = TRUE)

```

Jaccard index

```{r,jaccard}

jaccard <- function(a, b) {
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    return (intersection/union)
}

ora <- c(ora_ups,ora_dns)
fcs <- c(fgsea_ups,fgsea_dns)

jaccard(ora,fcs)

```

## Drill in to a specific pathway

In this case the top upregulated and downregulated pathways

```{r,eplot}

upname <- head(fgsea_up,1)$pathway
plotEnrichment(gsets[[upname]], stat, gseaParam = 1, ticksSize = 0.2)

dnname <- head(fgsea_dn,1)$pathway
plotEnrichment(gsets[[dnname]], stat, gseaParam = 1, ticksSize = 0.2)

```

Now a heatmap of these gene sets.

```{r,heat}

# reads per million normalisation
rpm <- apply(mxf,2, function(x) { x/sum(x) * 1000000} )
colnames(rpm) <- c("UNTR1","UNTR2","UNTR3","TRT1","TRT2","TRT3")
gnames_up <- gsets[[which(names(gsets) == upname)]]
gnames_dn <- gsets[[which(names(gsets) == dnname)]]
gene_ids <- rownames(mxf)
gene_names <- sapply(strsplit(gene_ids," "),"[[",2)
rpm_up <- rpm[which(gene_names %in% gnames_up),]
rownames(rpm_up) <- sapply(strsplit(rownames(rpm_up)," "),"[[",2)
rpm_dn <- rpm[which(gene_names %in% gnames_dn),]
rownames(rpm_dn) <- sapply(strsplit(rownames(rpm_dn)," "),"[[",2)
colsidecols <- c("blue","blue","blue","red","red","red")
heatmap.2(rpm_up,scale="row",trace="none",margins=c(10,15),main=upname,ColSideColors=colsidecols)
heatmap.2(rpm_dn,scale="row",trace="none",margins=c(10,15),main=dnname,ColSideColors=colsidecols)

```

## Session information

For reproducibility

<br><details><summary><b>Click HERE to show session info</b></summary><br><p>

```{r,session}

sessionInfo()

```

</details>