### 2023-06-30 example2.Rmd
code edits

---
title: "A second example of extremely reproducible enrichment analysis"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/enrichment_recipe

## Introduction

This guide is a Rmarkdown script that conducts differential expression and enrichment analysis,
which are very popular workflows for transcriptome data.

This is meant to be a boilerplate template, which you can remix and modify to suit your analytical
needs.

In the code chunk below called `libs`, you can add and remove required R library dependancies.
Check that the libraries listed here match the Dockerfile, otherwise you might get errors.

```{r,libs}

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("clusterProfiler")
  library("fgsea")
  library("eulerr")
  library("gplots")
})

```

For this guide I will be using bulk RNA-seq data from a previous study, which is
deposited at NCBI GEO with [accession GSE93236](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE93236).

The experiment is designed to investigate the effect of knowckdown of the histone methyltransferase set7.

We will be fetching the gene expression count matrix directly from NCBI GEO for downstream analysis.
Link: https://ftp.ncbi.nlm.nih.gov/geo/series/GSE93nnn/GSE93236/suppl/GSE93236_Keating_Set7_KD_counts.tsv.gz

```{r,fetch}

download.file("https://ftp.ncbi.nlm.nih.gov/geo/series/GSE93nnn/GSE93236/suppl/GSE93236_Keating_Set7_KD_counts.tsv.gz",
  destfile="GSE93236_Keating_Set7_KD_counts.tsv.gz")

x <- read.table("GSE93236_Keating_Set7_KD_counts.tsv.gz")

dim(x)

# there is a row we need to remove
mx <- x[rownames(x)!="NOTINBED",]

head(mx)

dim(mx)

```

The row names contains an underscore that we need to remove.

```{r,geneid_fix}

rownames(mx) <- sub("_"," ",rownames(mx))

head(mx)

```

Curate sample sheet from the colnames.

```{r,samplesheet1}

ss <- as.data.frame(colnames(x))
ss$trt <- as.numeric(grepl("Set7KD",colnames(x)))

ss %>%
  kbl(caption="sample sheet for Set7KD experiment") %>%
  kable_paper("hover", full_width = F)

```

## Data quality control

QC is important, even if you are using public transcriptome data.
For RNA-seq it is a good idea to show the number of reads for each sample.

```{r,counts1}

par(mar=c(5,7,5,1))
barplot(rev(colSums(mx)),horiz=TRUE,las=1,main="number of reads per sample in SRP038101")

```

Now make a MDS plot.

```{r,mds1}

mds <- cmdscale(dist(t(mx)))

# expand plot area
XMIN=min(mds[,1])*1.3
XMAX=max(mds[,1])*1.3
YMIN=min(mds[,2])*1.3
YMAX=max(mds[,2])*1.3

cols <- as.character(ss$trt)
cols <- gsub("0","lightblue",cols)
cols <- gsub("1","pink",cols)
plot(mds, xlab="Coordinate 1", ylab="Coordinate 2",
  xlim=c(XMIN,XMAX),ylim=c(YMIN,YMAX),
  pch=19,cex=2,col=cols, main="MDS plot")
text(cmdscale(dist(t(mx))), labels=colnames(mx) )

```

## Differential expression analysis

I will be using DESeq2 for DE analysis, the count matrix is prefiltered using
a detection threshold of 10 reads per sample across all samples.
All genes that meet the detection threshold will comprise the background list.
The first 6 rows of the count matrix are shows.

```{r,de1}

mxf <- mx[which(rowMeans(mx)>=10),]
dim(mxf)

head(mxf,6) %>%
  kbl(caption="Count matrix format") %>%
  kable_paper("hover", full_width = F)

dds <- DESeqDataSetFromMatrix(countData = mxf , colData = ss, design = ~ trt )
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <-cbind(as.data.frame(z),mxf)
def <-as.data.frame(zz[order(zz$pvalue),])

head(def,10) %>%
  kbl(caption="Top DE genes for Aza treatment") %>%
  kable_paper("hover", full_width = F)

```

Make a smear plot.

```{r,smearplot}

sigf <- subset(def,padj<=0.05)

DET=nrow(mxf)
NSIG=nrow(sigf)
NUP=nrow(subset(sigf,log2FoldChange>0))
NDN=nrow(subset(sigf,log2FoldChange<0))

HEADER=paste(DET,"detected genes;",NSIG,"w/FDR<0.05;",NUP,"up;",NDN,"down")

plot(log10(def$baseMean) ,def$log2FoldChange,
  cex=0.6,pch=19,col="darkgray",
  main="Set7KD in endothelial cells",
  xlab="log10(basemean)",ylab="log2(fold change)")

points(log10(sigf$baseMean) ,sigf$log2FoldChange,
  cex=0.6,pch=19,col="red")

mtext(HEADER)

```

In the next sections I will run enrichment analysis with over-representation test
and compare it to functional class scoring.

## Enrichment using over-representation test

I've compiled a reporting checklist:

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | GEO under accession GSE93236 |
| Other parameters | Min gene set size of 10 |

Get the gene sets loaded in R.
They are located in the `/ref` folder of the repo.

```{r,gs}
# from https://reactome.org/download/current/ReactomePathways.gmt.zip
genesets2 <- read.gmt("ref/ReactomePathways_2023-03-06.gmt")
gsets <- gmtPathways("ref/ReactomePathways_2023-03-06.gmt")

```

Now filter the gene names into three lists, up-regulated, down-regulated and
background.
Background is simply all genes that were detected.

```{r,subset1}

defup <- rownames(subset(def,padj<=0.05 & log2FoldChange>0))
defup <- unique(sapply(strsplit(defup," "),"[[",2))

defdn <- rownames(subset(def,padj<=0.05 & log2FoldChange<0))
defdn <- unique(sapply(strsplit(defdn," "),"[[",2))

bg <- rownames(def)
bg <- unique(sapply(strsplit(bg," "),"[[",2))

message("number of genes in each group")
lapply(list("background"=bg,"up-regulated"=defup,"down-regulated"=defdn),length)

```

One of the quirky things I learned about clusterprofiler is that
not all genes specified by the user as background are included in
the enrichmentcalculation.
It appears as though the genes not included in the annotation set are
discarded.
I believe this is a subtle but significant problem.
To fix it is relatively easy from the user's point of view.
Just create a new gene set that consists of all detected genes and add it
to the main gene set data frame.

Adding all detected genes to the background appears to improve results!

```{r,gs2}

bgdf <- data.frame("background",bg)
colnames(bgdf) <- c("term","gene")
genesets2 <- rbind(genesets2,bgdf)

```

Enrichment firstly with upregulated genes.
Note that the BgRatio contains the correct number of genes in the background now.

```{r,oraup}

# show 10 pathways only
n_pw=10

ora_up <- as.data.frame(enricher(gene = defup ,
  universe = bg,  maxGSSize = 50000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora_up$geneID <- NULL
ora_up <- subset(ora_up,p.adjust<0.05 & Count >=10)
ora_ups <- rownames(ora_up)

gr <- as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",2))

ora_up$es <- gr/br
ora_up <- ora_up[order(-ora_up$es),]
ora_up$Description=NULL

head(ora_up,n_pw) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

topup2 <- rev(head(ora_up$es,10))
names(topup2) <- rev(head(ora_up$ID,10))

```

Now with the downregulated genes

```{r,oradn}

ora_dn <- as.data.frame(enricher(gene = defdn ,
  universe = bg,  maxGSSize = 50000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora_dn$geneID <- NULL
ora_dn <- subset(ora_dn,p.adjust<0.05 & Count >=10)
ora_dns <- rownames(ora_dn)

gr <- as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",2))

ora_dn$es <- gr/br
ora_dn <- ora_dn[order(-ora_dn$es),]
ora_dn$Description=NULL

head(ora_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

topdn2 <- head(ora_dn$es,n_pw)
names(topdn2) <- head(ora_dn$ID,n_pw)

```

Make a barplot

```{r,topbarplot2}

par(mar=c(5,20,5,1))

cols <- c(rep("blue",n_pw),rep("red",n_pw))

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE Reactomes",
  xlab="ES",
  cex.main=0.9)

mtext("ORA test")

```

## Enrichment analysis with functional class scoring

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | FGSEA (check version at foot of report) |
| Statistical test used | Kolmogorov-Smirnov test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | GEO under accession GSE93236 |
| Other parameters | Min gene set size of 10 |

```{r,fgsea}

def2 <- def
def2$genename <- sapply(strsplit(rownames(def2)," "),"[[",2)
def2 <- def2[,c("stat","genename")]
def2 <- aggregate(. ~ genename, def2, sum)
stat <- def2$stat
names(stat) <- def2$genename
fgseaRes <- fgsea(pathways=gsets, stats=stat, minSize=10, nPermSimple = 10000)
fgseaRes <- fgseaRes[order(fgseaRes$pval),]

```

```{r,fgsea_sig}

fgsea_up <- subset(fgseaRes,padj<0.05 & ES>0)
fgsea_ups <- fgsea_up$pathway
fgsea_dn <- subset(fgseaRes,padj<0.05 & ES<0)
fgsea_dns <- fgsea_dn$pathway

fgsea_up <-  data.frame(fgsea_up[order(-fgsea_up$ES),])

fgsea_dn <-  data.frame(fgsea_dn[order(fgsea_dn$ES),])

head(fgsea_up,n_pw) %>%
  kbl(row.names = FALSE, caption="FGSEA:Top upregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

head(fgsea_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="FGSEA:Top downregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

fgsea_up <- head(fgsea_up,n_pw)
fgsea_up_vec <- fgsea_up$ES
names(fgsea_up_vec) <- fgsea_up$pathway
fgsea_up_vec <- rev(fgsea_up_vec)

fgsea_dn <- head(fgsea_dn,n_pw)
fgsea_dn_vec <- fgsea_dn$ES * -1
names(fgsea_dn_vec) <- fgsea_dn$pathway

```

Make a barplot

```{r,topbarplot3}

par(mar=c(5,20,5,1))

cols <- c(rep("blue",n_pw),rep("red",n_pw))

barplot(c(fgsea_dn_vec,fgsea_up_vec),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE Reactomes",
  xlab="ES",
  cex.main=0.9)

mtext("FCS test")

```

## Compare pathway results

```{r,venn1,fig.width=8,fig.height=5}

v1 <- list("ORA up"=ora_ups, "FCS up"=fgsea_ups)
v2 <- list("ORA dn"=ora_dns, "FCS dn"=fgsea_dns)
par(mar=c(10,10,10,10))
par(mfrow=c(2,1))
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2))
plot(euler(v2),quantities = list(cex = 2), labels = list(cex = 2))

```

Jaccard index

```{r,jaccard}

jaccard <- function(a, b) {
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    return (intersection/union)
}

ora <- c(ora_ups,ora_dns)
fcs <- c(fgsea_ups,fgsea_dns)

jaccard(ora,fcs)

```

## Drill in to a specific pathway

In this case the top upregulated and downregulated pathways

```{r,eplot}

upname <- head(fgsea_up,1)$pathway
plotEnrichment(gsets[[upname]], stat, gseaParam = 1, ticksSize = 0.2)

dnname <- head(fgsea_dn,1)$pathway
plotEnrichment(gsets[[dnname]], stat, gseaParam = 1, ticksSize = 0.2)

```

Now a heatmap of these gene sets.

```{r,heat}

# reads per million normalisation
rpm <- apply(mxf,2, function(x) { x/sum(x) * 1000000} )
colnames(rpm) <- c("UNTR1","UNTR2","UNTR3","TRT1","TRT2","TRT3")
gnames_up <- gsets[[which(names(gsets) == upname)]]
gnames_dn <- gsets[[which(names(gsets) == dnname)]]
gene_ids <- rownames(mxf)
gene_names <- sapply(strsplit(gene_ids," "),"[[",2)
rpm_up <- rpm[which(gene_names %in% gnames_up),]
rownames(rpm_up) <- sapply(strsplit(rownames(rpm_up)," "),"[[",2)
rpm_dn <- rpm[which(gene_names %in% gnames_dn),]
rownames(rpm_dn) <- sapply(strsplit(rownames(rpm_dn)," "),"[[",2)
colsidecols <- c("blue","blue","blue","red","red","red")
heatmap.2(rpm_up,scale="row",trace="none",margins=c(10,15),main=upname,ColSideColors=colsidecols)
heatmap.2(rpm_dn,scale="row",trace="none",margins=c(10,15),main=dnname,ColSideColors=colsidecols)

```

## Session information

For reproducibility

<br><details><summary><b>Click HERE to show session info</b></summary><br><p>

```{r,session}

sessionInfo()

```

</details>