### 2023-12-13 manuscript/letter.Rmd
2013 data

---
title: "How vanishing web-based tools hamper reproducibility"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
bibliography: references.bib
csl: genome-biology.csl
---

[Anusuiya Bora<sup>1,2</sup>](https://orcid.org/0009-0006-2908-1352),
[Nhi Nguyen<sup>2</sup>](https://orcid.org/XXXX-XXXX-XXXX-XXXX),
[Mark Ziemann<sup>1,2</sup>*](https://orcid.org/0000-0002-7688-6974)

**Affiliations**

1. Burnet Institute, Melbourne, Australia.

2. Deakin University, School of Life and Environmental Sciences, Geelong, Australia.

(*) Corresponding author: mark.ziemann@burnet.edu.au

## Abstract

Web-based tools are an indispensable resource for conducting genomic and bioinformatics analysis, but if they disappear
it can leave reproducibility gaps in research projects.
Unless such tools allow for preservation of key materials for reproduction, researchers may unwittingly be in breach of
data retention mandates.

## Main text

According to a 2016 survey, 52% of researchers believe we are in a reproducibility crisis [@Baker2016-hm].
In genomics we don't have a lot of data on study reproducibility, but a small survey from 2009 indicates that most
studies are irreproducible due to a lack of shared data, software and documentation [@Ioannidis2009-hv].
The types of tools we use for the analysis also play a role.
Web-based point-and-click tools have become very popular in genomics for the analysis of gene lists, to
interpret whether certain biochemical or signaling pathways are over-represented in a particular genomic study.
These tools are so popular because they do not require the installation of any software, and a gene list analysis
can be conducted in a matter of seconds.
They also require little training to use, which on one hand simplifies bioinformatics tasks, but on the other hand
these tools can easily be misused and results misreported, which can lead to misleading conclusions
[ @Timmons2015-ex ; @Wijesooriya2022-li].
Last year we sought to test whether a small group of articles published in 2019 involving enrichment analysis with the DAVID tool [@Sherman2022-hf]
were reproducible using the authors' own methods.
We found only 4/20 enrichment studies had a high degree of reproduction, while 7/20 had severe discrepancies.
After our pilot study completed we were surprised to hear that the version of DAVID used for all of these studies (v6.8) would no longer be available from June 2022 onwards.
This was disturbing, as according to analysis of PubMed citations, some 20,000 articles citing DAVID enrichment tools will no longer be reproducible with
the original tools, some articles just one or two years old.
This is a prime example of "link decay", a phenomenon where internet based resources are lost over time, which has been raised as a significant and ongoing problem for bioinformatic reproducibility [@Hennessey2013-tj].
Of course DAVID isn't the only webserver without the option of preservation; our analysis of the latest NAR Webserver issue shows 28% (23/81) of webserver tools lacked source code or other features to allow preservation [@Seelow2023-xk].
A look at NAR Webservers from 2013 shows 68% of tools did not provide any preservation options, and 35% (33/95) are no longer online.

Researchers should keep in mind that institutional and funding mandates for data retention also apply to software and algorithms [@NHMRC2019-wu ; @Harvard_2023-cl].
Therefore, to remain compliant with these mandates and support reproducibility, we caution against using data from web-based tools for publication unless there is a system in place to enable future reproducibility.
Scripted workflows are perhaps the most suitable solution for enabling reproduction in the long-term, but require expertise in computer programming.
We have written a protocol for extremely reproducible enrichment analysis in R, designed for novice bioinformaticians to fill in this gap [@Ziemann2023-kz], but we understand this isn't for everyone.
ShinyGO is another potential remedy, as it provides most of the necessary features in a point-and-click web interface but also allows users to download all historical versions of the tool to be run as
a Docker based web page on the users own hardware [@Ge2020-qe].
In order to get genomics research to meet the ten-year reproducibility challenge [@Perkel2020-io], we need progress towards better tooling, a greater emphasis on investigator training, better institutional support,
clearer funder mandates, stronger publication criteria, and more incentives for rigourous and reliable practices [@Lewis2016-qh; @Diong2021-ma].

## Competing Interests

No competing interests were disclosed.

## Acknowledgements

We thank Ms Claudia Beyer for advice.

## Bibliography