### 2025-08-29 example_10m.Rmd
using data to support the 10 mistakes article

---
title: "Supporting 10 common mistakes article"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/enrichment_recipe

## Introduction

```{r,libs}

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("clusterProfiler")
  library("eulerr")
  library("gplots")
  library("mitch")
  library("fgsea")
  library("RhpcBLASctl")
})

RhpcBLASctl::blas_set_num_threads(1)

```

For this guide I will be using bulk RNA-seq data from a previous study, which is
deposited at NCBI GEO and SRA under accession numbers: GSE55123 and SRP038101 (Lund et al, 2014).
The experiment is designed to investigate the effect of Azacitidine treatment on AML3 cells.

The raw data have been processed by the DEE2 project, and the summary gene expression counts are
available at the dee2.io website, and programmatically with the getDEE2 bioconductor package
(Ziemann et al, 2019).

The gene counts have also been deposited to the `/data` folder in the `example.Rdata` file in case
the DEE2 resource becomes unavailable.
To import it, use the command: `load("data/example.Rdata")`

Alternatively, you could fetch data from another resource like NCBI GEO, Zenodo or from the host
storage drive.

```{r,fetch}

mdat <- getDEE2Metadata("hsapiens")

# get sample sheet
ss <- subset(mdat,SRP_accession=="SRP038101")

# fetch the whole set of RNA-seq data
x <- getDEE2("hsapiens",ss$SRR_accession , metadata=mdat, legacy=TRUE)
mx <- x$GeneCounts
rownames(mx) <- paste(rownames(mx),x$GeneInfo$GeneSymbol)
dim(mx)

# aza no filtering
ss$trt <- grepl("Treated",ss$Experiment_title)

ss %>%
  kbl(caption="sample sheet for Aza treatment in AML3 cells") %>%
  kable_paper("hover", full_width = F)

```

## Data quality control

QC is important, even if you are using public transcriptome data.
For RNA-seq it is a good idea to show the number of reads for each sample.

```{r,counts1}

par(mar=c(5,7,5,1))
barplot(rev(colSums(mx)),horiz=TRUE,las=1,main="number of reads per sample in SRP038101")

```

Now make a MDS plot.

```{r,mds1}

mds <- cmdscale(dist(t(mx)))

# expand plot area
XMIN=min(mds[,1])*1.3
XMAX=max(mds[,1])*1.3
YMIN=min(mds[,2])*1.3
YMAX=max(mds[,2])*1.3

cols <- as.character(grepl("Treated",ss$Experiment_title))
cols <- gsub("FALSE","lightblue",cols)
cols <- gsub("TRUE","pink",cols)
plot(mds, xlab="Coordinate 1", ylab="Coordinate 2",
  xlim=c(XMIN,XMAX),ylim=c(YMIN,YMAX),
  pch=19,cex=2,col=cols, main="MDS plot")
text(cmdscale(dist(t(mx))), labels=colnames(mx) )

```

## Differential expression analysis

I will be using DESeq2 for DE analysis, the count matrix is prefiltered using
a detection threshold of 10 reads per sample across all samples.
All genes that meet the detection threshold will comprise the background list.
The first 6 rows of the count matrix are shows.

```{r,de1}

mxf <- mx[which(rowMeans(mx)>=10),]
dim(mxf)

head(mxf,6) %>%
  kbl(caption="Count matrix format") %>%
  kable_paper("hover", full_width = F)

dds <- DESeqDataSetFromMatrix(countData = mxf , colData = ss, design = ~ trt )
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <-cbind(as.data.frame(z),mxf)
def <-as.data.frame(zz[order(zz$pvalue),])

head(def,10) %>%
  kbl(caption="Top DE genes for Aza treatment") %>%
  kable_paper("hover", full_width = F)

```

Make a smear plot.

```{r,smearplot}

sigf <- subset(def,padj<=0.05)

DET=nrow(mxf)
NSIG=nrow(sigf)
NUP=nrow(subset(sigf,log2FoldChange>0))
NDN=nrow(subset(sigf,log2FoldChange<0))

HEADER=paste(DET,"detected genes;",NSIG,"w/FDR<0.05;",NUP,"up;",NDN,"down")

plot(log10(def$baseMean) ,def$log2FoldChange,
  cex=0.6,pch=19,col="darkgray",
  main="5-azacitidine treatment in AML3 cells",
  xlab="log10(basemean)",ylab="log2(fold change)")

points(log10(sigf$baseMean) ,sigf$log2FoldChange,
  cex=0.6,pch=19,col="red")

mtext(HEADER)

```

In the next sections I will run enrichment analysis with over-representation test
and compare it to functional class scoring.

## Get gene sets

```{r,gs}
# from https://reactome.org/download/current/ReactomePathways.gmt.zip
genesets2 <- read.gmt("ref/ReactomePathways_2023-03-06.gmt")
gsets <- gmtPathways("ref/ReactomePathways_2023-03-06.gmt")

```

## Mistake 1 - Raw p-values

## Mistake 2 - Background

## Mistake 4 - prioritisation by p-values

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | FORA (check fgsea version at foot of report) |
| Statistical test used | hypergeometric |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection | DESeq2 FDR<0.05 and log2FC>0 or log2FC<0 |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 5 |

```{r,mistake2}

up <- unique(sapply(strsplit(rownames(subset(def, log2FoldChange>0 & padj<0.05))," "),"[[",2))

dn <- unique(sapply(strsplit(rownames(subset(def, log2FoldChange<0 & padj<0.05))," "),"[[",2))

bg <- unique(sapply(strsplit(rownames(def)," "),"[[",2))

gt <- read.table("https://dee2.io/data/hsapiens/hsa_gene_info.tsv",header=TRUE)
head(gt)

all <- unique(gt$GeneSymbol)

lapply(list(up,dn,bg,all),length)

fres_up_bg <- fora(pathways=gsets, genes=up, universe=bg, minSize = 5, maxSize = Inf)
fres_up_bg <- subset(fres_up_bg,padj<0.05)$pathway
length(fres_up_bg)

fres_up_all <- fora(pathways=gsets, genes=up, universe=all, minSize = 5, maxSize = Inf)
fres_up_all <- subset(fres_up_all,padj<0.05)$pathway
length(fres_up_all)

fres_dn_bg <- fora(pathways=gsets, genes=dn, universe=bg, minSize = 5, maxSize = Inf)
fres_dn_bg <- subset(fres_dn_bg,padj<0.05)$pathway
length(fres_dn_bg)

fres_dn_all <- fora(pathways=gsets, genes=dn, universe=all, minSize = 5, maxSize = Inf)
fres_dn_all <- subset(fres_dn_all,padj<0.05)$pathway
length(fres_dn_all)

v1 <- list("up"=fres_up_bg, "up*"=fres_up_all, "dn"=fres_dn_bg, "dn*"=fres_dn_all)

par(mar=c(1,1,1,1))
par(mfrow=c(1,1))
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2))

v2 <- list("up"=fres_up_bg, "up*"=fres_up_all)
plot(euler(v2),quantities = list(cex = 2), labels = list(cex = 2))

par(mar=c(5.1,4.1,4.1,2.1))

```

## Mistake 4 - prioritisation by p-values

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | mitch (check version at foot of report) |
| Statistical test used | Rank ANOVA (Kruskal Wallis) |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Scoring | DESeq2 Wald stat |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 5 |

```{r,mistake4a}

gt <- def
gt$genename <- sapply(strsplit(rownames(gt)," "),"[[",2)
gt <- gt[,"genename",drop=FALSE]
gt$id <- rownames(gt)
m <- mitch_import(x=def,DEtype="deseq2",geneTable=gt)
mres <- mitch_calc(x=m,genesets=gsets,minsetsize=5,cores=4,priority="significance")

mres_sig_up <- head(subset(mres$enrichment_result,s.dist>0),10)
mres_sig_dn <- head(subset(mres$enrichment_result,s.dist<0),10)

mres_sig_up %>%
  kbl(caption="Top upregulated pathways when prioritised by FDR") %>%
  kable_paper("hover", full_width = F)

mres_sig_dn %>%
  kbl(caption="Top downregulated pathways when prioritised by FDR") %>%
  kable_paper("hover", full_width = F)

mres <- mitch_calc(x=m,genesets=gsets,minsetsize=5,cores=4,priority="effect")

mres_es_up <- head(subset(mres$enrichment_result,s.dist>0 & p.adjustANOVA < 0.05),10)
mres_es_dn <- head(subset(mres$enrichment_result,s.dist<0 & p.adjustANOVA < 0.05),10)

mres_es_up %>%
  kbl(caption="Top upregulated pathways when prioritised by ES (discarding any with FDR>0.05)") %>%
  kable_paper("hover", full_width = F)

mres_es_dn %>%
  kbl(caption="Top upregulated pathways when prioritised by ES (discarding any with FDR>0.05)") %>%
  kable_paper("hover", full_width = F)

```

Make a barplot

```{r,mistake4b,fig.height=6,fig.width=6}

par(mar=c(2,20,2,0))

mres_sig_up_vals <- -log(mres_sig_up$pANOVA)
names(mres_sig_up_vals) <- mres_sig_up$set

mres_sig_dn_vals <- -log(mres_sig_dn$pANOVA)
names(mres_sig_dn_vals) <- mres_sig_dn$set

mres_es_up_vals <- mres_es_up$s.dist
names(mres_es_up_vals) <- mres_es_up$set

mres_es_dn_vals <- -1 * mres_es_dn$s.dist
names(mres_es_dn_vals) <- mres_es_dn$set

par(mfrow=c(2,1))

barplot( mres_sig_up_vals,
  horiz=TRUE,las=1,cex.names=0.65,col="gray",
  main="Up Sig (-log10 p-value)",
  xlab="-log10 p-value",
  cex.main=0.9)

barplot( mres_es_up_vals,
  horiz=TRUE,las=1,cex.names=0.65,col="gray",
  main="Up ES",
  xlab="ES",
  cex.main=0.9)

barplot( mres_sig_dn_vals,
  horiz=TRUE,las=1,cex.names=0.65,col="gray",
  main="Down Sig (-log10 p-value)",
  xlab="-log10 p-value",
  cex.main=0.9)

barplot( mres_es_dn_vals,
  horiz=TRUE,las=1,cex.names=0.65,col="gray",
  main="Down ES",
  xlab="-ES",
  cex.main=0.9)

par(mfrow=c(1,1))
par(mar=c(5.1,4.1,4.1,2.1))

```

## Mistake 5 - gene set list size too small or long

```{r,mistake5}

bg <- unique(sapply(strsplit(rownames(def)," "),"[[",2))

nrow(subset(def,padj<0.05 & log2FoldChange>0))
nrow(subset(def,padj<0.05 & log2FoldChange<0))
nrow(def)

sizes <- c(50,100,200,300,400,500,750,1000,1500,2000,2500,3000,4000,5000,6000,7000)

fup <- nrow(subset(fora(pathways=gsets, genes=up, universe=bg, minSize = 5, maxSize = Inf),padj<0.05))
fdn <- nrow(subset(fora(pathways=gsets, genes=dn, universe=bg, minSize = 5, maxSize = Inf),padj<0.05))

nups <- unlist(lapply(sizes, function(i) {
  genes <- unique(sapply(strsplit(rownames(head(subset(def, log2FoldChange>0),i))," "),"[[",2))
  fres <- fora(pathways=gsets, genes=genes, universe=bg, minSize = 5, maxSize = Inf)
  nrow(subset(fres,padj<0.05))
}))

ndns <- unlist(lapply(sizes, function(i) {
  genes <- unique(sapply(strsplit(rownames(head(subset(def, log2FoldChange<0),i))," "),"[[",2))
  fres <- fora(pathways=gsets, genes=genes, universe=bg, minSize = 5, maxSize = Inf)
  nrow(subset(fres,padj<0.05))
}))

df <- data.frame("up"=nups,"dn"=ndns,row.names=sizes)
df
plot(sizes,df$up,type="b",col="red",ylab="no. gene sets with FDR<0.05",xlab="gene list length")
lines(sizes,df$dn,type="b",col="blue")

plot(sizes,df$up,type="b",col="red",ylab="no. gene sets with FDR<0.05",xlab="gene list length",xlim=c(0,500))
lines(sizes,df$dn,type="b",col="blue")

```

## Session information

For reproducibility

<br><details><summary><b>Click HERE to show session info</b></summary><br><p>

```{r,session}

save.image("example_10m.Rdata")
sessionInfo()

```

</details>